Report on Late Winter2013 Production: Image Differencing\vspace{-2ex}

### 2013-03-29 W13report.tex
Wrote summary section.

\author{Andrew Becker, Simon Krughoff, Andrew Connolly, Russell Owen}
\title{Report on Late Winter2013 Production: Image Differencing}
\date{\today}

@@

\maketitle

{\bf WRITE ME}

>>>>

\author{Andrew Becker, Simon Krughoff, Andrew Connolly, Russell Owen\vspace{-2ex}}
\title{Report on Late Winter2013 Production: Image Differencing\vspace{-2ex}}
\date{\vspace{-2ex}\today}

@@

\maketitle
\vspace{-2ex}
We provide here a technical report on the LSST Data Management
late--Winter 2013 production runs.  This production focused on the
evaluation of LSST DM image differencing code, in particular on
evaluating the rate of false detections the pipeline generates.  As
inputs, we used a suite of simulated images generated in idealized
conditions and containing no intrinsic variability.  Thus {\it any}
sources detected in these image differences are ``false positives''.

We investigated two distinct operational modes of the pipeline.  The
first pre--filters the science image with its own PSF, yielding a
point--source maximum likelihood image.  The template is then
PSF--matched to this image, subtracted from it, and detection and
measurement happen directly on this difference.  The second mode
matches the template image to the science image, creates their
difference, post--filters the difference with the science image PSF,
and then runs detection on this maximum likelihood image.  Measurement
happens on the detected footprints, but in the difference image.  We
also investigated the complexity of the basis functions used to model
the PSF--matching kernel, and the degree to which this kernel varies
across the images.  We generated several goodness of fit statistics
for a control sample of sources that were {\it not} used in the kernel
fit, in order to assess the interpolation ability of the matching
kernel.

We compared the measured numbers of false detections to the number we
expect due to random fluctuations in a background Gaussian field.  The
latter is a function of both seeing in the images, and of the
detection threshold used.  At the fiducial detection threshold of
5--sigma, the number of expected false positives per LSST sensor is of
order {\bf XX}, and we measure of order {\bf YY}, indicating that we
are able to achieve the systematic floor in false detection rate.  We
examined our measured false detection rate as a function of detection
threshold and find that it roughly scales as expected.  In order for
there to be less than 1 expected false detection per sensor, we must
operate at a detection threshold of {\bf X.X} sigma.

We found a severe dependence of the false positive rate on our
understanding of the variance; even a 1\% misestimate of the noise may
increase or decrease the false positive rate by a factor of {\bf XX}
at 5--sigma.  Prefiltering is preferred in this regard, as our
variance planes are shown to track the variance to higher fidelity at
the detection stage.  We find that over/undersubtraction of the image
background leads to a bias in the positive to negative false detection
ratio, but not on the overall numbers of detections.  Prefiltering
also prevents, in many cases, a deconvolution of the template image
when the science image has better seeing, which is a clear advantage
of the technique.

We outline several additional steps we will pursue on the Summer 2013
timescale, including operating on images with more astrophysical and
atmospheric complexity, improving measurement on PSF--filtered
sources, adopting Gaussian processes for modeling of the spatial
kernel, and using statistics calculated during the image subtraction
to tune the algorithm ``on the fly'' (i.e. predict the numbers of
false positives it will generate).
